---
title: "Final Project"
author:
- Winnie
- June
- Helen
- Akshaj
- Lilly
output:
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=5, warning=FALSE, message=FALSE)
options(scipen = 0, digits = 3)
```


```{r load-packages}
# Load packages
if(!require('pacman')) install.packages('pacman')
pacman::p_load(tidyverse, readxl, skimr, corrplot, GGally, gridExtra, rpart, rpart.plot, caret, vip, randomForest, e1071)
```


**1. Goal of study**. Restate the goal of study and importance of your project. 

The goal of our project is to analyze the predictors that lead to greater poverty rates. Our project’s model is based on the idea that factors such as Education costs, per capita personal income, population density, unemployment rate and more, determine poverty. Our regression model, along with our analysis, should give the class and our team members a much greater understanding of what factors impact the different communities around the world which suffer from substantial impoverished conditions, ultimately leading us to draw conclusions based upon our findings. 


**2. Data:**

i) What is your data source?

Our data sources are from US department of Labor, US Bureau of Labor Statistics, US Census Bureau, and other official government sources.

ii) Describe how you acquire the data. (You can provide another set of code).

We manually downloaded datasources to assemble a singular dataset for my needs.

iii) Describe how you clean the data (e.g. missing value, errors, outliners).

We didn't have much to do with cleaning the data as we used a standardized government format/official format to collect our data, so there were wasn't much to clean.

iv) Is your data representative enough for your goal of study? Could there be any bias? How will you address the bias if any?

Yes, our data is representative enough due to the wide range of data that we are using in our study. However, aas all of these data sources are from the US government, there might be some bias when it comes to our data.


**3. EDA**

As the first step of our analysis, we need to make sense out of the data by EDA. 
Please provide a complete EDA of your data using relevant plots (with title, labels, colors, legends) and descriptions.
What have you found so far? 


# Research Question


**What state-level variables most strongly predict higher poverty rates across the United States?**


- Unit of analysis: 50 U.S. states
- Time period: 2023
- Outcome: Poverty rate (%)
- Predictors: Economic, social, and policy variables


---


# Load Data


```{r load-data}
library(dplyr)

# Load your cleaned dataset
# Update the file path to match your data location
state_data <- read.csv("/Users/winniex/Desktop/WDS_2026_AI_HOME/Project/dataset - Sheet1 (2).csv", sep=",", header=T, as.is=T)

state_data <- state_data %>%
  rename("state"="State")

# Show first few rows
head(state_data)
```



```{r data-structure}
# Data structure
glimpse(state_data)
```


---


# Data Quality Check
```{r}
str(state_data)
```

```{r}
summary(state_data)
```

```{r data-summary}
# Summary statistics for all variables
skim(state_data)
```


```{r missing-values}
# Check for missing values
missing_counts <- colSums(is.na(state_data))
missing_counts[missing_counts > 0]
```


```{r basic-stats}
# Basic dataset info
cat("Number of states:", nrow(state_data), "\n")
cat("Number of variables:", ncol(state_data), "\n")
```

---

# Outcome Variable: Poverty Rate


```{r poverty-summary}
# Summary statistics
summary(state_data$poverty)
```


```{r poverty-distribution}
# Histogram
ggplot(state_data, aes(x = poverty)) +
 geom_histogram(bins = 10, fill = "steelblue", alpha = 0.7) +
 geom_vline(xintercept = mean(state_data$poverty),
            color = "red", linetype = "dashed", size = 1) +
 labs(title = "Distribution of Poverty Rates Across States",
      subtitle = "Red line = mean",
      x = "Poverty Rate (%)",
      y = "Count") +
 theme_minimal()
```

```{r}
library(usmap)

stat.poverty.plot <- plot_usmap(regions = "state",                   
    data = state_data, 
    values = "poverty", color = "black") + 
    labs(title = "State Poverty Rate", subtitle = "Continental US States") +
    theme(legend.position = "right")

stat.poverty.plot
```

---

# Correlation Analysis

```{r}
# Select only numeric variables
numeric_vars <- state_data %>%
 select(where(is.numeric)) 

# Calculate correlation matrix
cor_matrix <- cor(numeric_vars, use = "complete.obs")

# Visualize correlation matrix
corrplot(cor_matrix,
        method = "color",
        type = "upper",
        tl.col = "black",
        title = "Correlation Matrix")
```

```{r}
# Extract correlations with poverty rate
cor_with_poverty <- cor_matrix["poverty", ] %>%
 sort(decreasing = TRUE)


# Display all correlations with poverty
cat("Correlations with poverty rate (sorted):\n")
print(round(cor_with_poverty, 3))
```



```{r correlation-table}
# Top 10 correlations (absolute value)
top_cors <- data.frame(
 Variable = names(cor_with_poverty)[-1],  # exclude poverty itself
 Correlation = cor_with_poverty[-1]
) %>%
 arrange(desc(abs(Correlation))) %>%
 slice(1:10)


cat("\nTop 10 predictors by absolute correlation:\n")
print(top_cors)
```

---

# Scatterplots: Top Predictors


```{r identify-top-predictors}
# Get top 4 predictors (by absolute correlation)
top_4_vars <- top_cors %>%
 slice(1:4) %>%
 pull(Variable)


cat("Top 4 predictors:\n")
print(top_4_vars)
```


```{r scatterplots}
# Create scatterplots for top 4 predictors
plot_list <- lapply(top_4_vars, function(var) {
 ggplot(state_data, aes_string(x = var, y = "poverty")) +
   geom_point(color = "steelblue", size = 2.5, alpha = 0.6) +
   geom_smooth(method = "lm", se = TRUE, color = "red", size = 1) +
   labs(title = paste("Poverty vs", var),
        x = var,
        y = "Poverty Rate (%)") +
   theme_minimal()
})

# Display in grid
grid.arrange(grobs = plot_list, ncol = 2)
```

---

# Pairs Plot


```{r pairs-plot}
# Select poverty rate + top 4 predictors
pairs_data <- state_data %>%
 select(poverty, all_of(top_4_vars))

# Create pairs plot
ggpairs(pairs_data,
       title = "Pairs Plot: Poverty Rate and Top 4 Predictors",
       lower = list(continuous = "smooth"),
       upper = list(continuous = "cor"))
```

---

# Distribution of Key Predictors

```{r predictor-distributions}
# Histograms for top 4 predictors
plot_list_hist <- lapply(top_4_vars, function(var) {
 ggplot(state_data, aes_string(x = var)) +
   geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
   labs(title = var, x = "", y = "Count") +
   theme_minimal()
})

# Display
grid.arrange(grobs = plot_list_hist, ncol = 2)
```

---

# Summary Statistics

```{r summary-stats}
# Summary for poverty rate and top predictors
summary_vars <- state_data %>%
 select(poverty, all_of(top_4_vars))

cat("Summary statistics for key variables:\n")
summary(summary_vars)
```


```{r correlation-strength}
# Categorize correlation strength
top_cors_categorized <- top_cors %>%
 mutate(
   Strength = case_when(
     abs(Correlation) > 0.7 ~ "Strong",
     abs(Correlation) > 0.4 ~ "Moderate",
     abs(Correlation) > 0.2 ~ "Weak",
     TRUE ~ "Very Weak"
   ),
   Direction = ifelse(Correlation > 0, "Positive", "Negative")
 )

cat("\nCorrelation strength and direction:\n")
print(top_cors_categorized)
```

---

# Key Findings

```{r key-findings}
# Calculate key metrics
mean_poverty <- mean(state_data$poverty)
sd_poverty <- sd(state_data$poverty)
min_poverty <- min(state_data$poverty)
max_poverty <- max(state_data$poverty)


cat("=== KEY FINDINGS ===\n\n")
cat("Dataset:\n")
cat("  - States:", nrow(state_data), "\n")
cat("  - Variables:", ncol(state_data), "\n\n")


cat("Poverty Rate:\n")
cat("  - Mean:", round(mean_poverty, 2), "%\n")
cat("  - SD:", round(sd_poverty, 2), "%\n")
cat("  - Range:", round(min_poverty, 2), "% to", round(max_poverty, 2), "%\n\n")


cat("Strongest Predictors:\n")
for(i in 1:4) {
 cat("  ", i, ".", top_cors_categorized$Variable[i],
     "(r =", round(top_cors_categorized$Correlation[i], 3), ")\n")
}


cat("\nNext Steps:\n")
cat("  - Build OLS regression model\n")
cat("  - Run LASSO feature selection\n")
cat("  - Check for multicollinearity (VIF)\n")
cat("  - Consider log transformations for skewed variables\n")
```


**4. Preliminary analysis**

With all the techniques we have learned so far, let's try to build models for your goal of study. 
Please summarize what you have learned from the model. Do the results make sense? 

# Load Data and Split into Train/Test Sets

```{r}
library(rpart)
library(rpart.plot)
library(caret)

set.seed(123)
train_index <- createDataPartition(state_data$poverty, p = 0.8, list = FALSE)
train_data <- state_data[train_index, ]
test_data <- state_data[-train_index, ]

train_data$state <- NULL
test_data$state  <- NULL
```

# Linear Regression

```{r}
fit1 <- lm(poverty ~ ., train_data)

summary(fit1)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(fit1, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

# Relex LASSO

```{r}
library(glmnet)

y <- train_data$poverty # extract Y
x <- model.matrix(poverty~., data=train_data)

set.seed(10)

lasso_cv <- cv.glmnet(x, y, alpha = 1, nfolds=10)

coef(lasso_cv, s = "lambda.1se")
```

```{r}
coef(lasso_cv, s = "lambda.min")
```

```{r}
plot(lasso_cv)
```

```{r}
library(coefplot)

#lambda.min
coef.min <- extract.coef(lasso_cv, lambda="lambda.min")
coef.min
```

```{r}
var_min <- rownames(coef.min)[-1]
min_data <- train_data[, c("poverty", var_min)]
min_data
```

```{r}
#lambda.1se
coef.1se <- extract.coef(lasso_cv, lambda="lambda.1se")
coef.1se
```

```{r}
var_1se <- rownames(coef.1se)[-1]
se_data <- train_data[, c("poverty", var_1se)]
se_data
```

```{r}
#min
fit.min <- lm(poverty ~ ., min_data)
summary(fit.min)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(fit.min, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

```{r}
#1se
fit.1se <- lm(poverty ~ ., se_data)
summary(fit.1se)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(fit.1se, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

# Decision Tree

```{r}
#Train the Model
tree_model <- rpart(poverty ~ ., 
                    data = train_data, 
                    method = "anova",  # for regression
                    control = rpart.control(minsplit = 10, cp = 0.01))

tree_model
```

```{r}
#Visualize the Tree
rpart.plot(tree_model, box.palette = "auto", nn = TRUE)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(tree_model, test_data, type = "vector")

# Evaluate Regression performance instead of using a Confusion Matrix
library(caret)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```


```{r}
#Prune the Tree
printcp(tree_model)  # Identify optimal cp (xerror is minimized)
optimal_cp <- tree_model$cptable[which.min(tree_model$cptable[, "xerror"]), "CP"]
pruned_tree <- prune(tree_model, cp = optimal_cp)
rpart.plot(pruned_tree)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(pruned_tree, test_data, type = "vector")

# Evaluate Regression performance instead of using a Confusion Matrix
library(caret)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

```{r}
# Tuning Hyperparameters
set.seed(10)

control <- trainControl(method = "cv", number = 10)  # 10-fold cross-validation
tuned_tree <- train(poverty ~ ., 
                    data = train_data, 
                    method = "rpart",
                    trControl = control,
                    tuneGrid = expand.grid(cp = seq(0.1, 0.6, by = 0.05)))  # Try cp values from 0 to 1 in increments of 0.1
print(tuned_tree)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(tuned_tree, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

```{r}
# 1. Extract the feature importance
# Decision trees store importance directly inside the model object
tree_importance <- tree_model$variable.importance
print(tree_importance)

# 2. Convert to a data frame and plot (same logic as before)
importance_df <- data.frame(
  Feature = names(tree_importance),
  Importance = as.numeric(tree_importance)
)

library(ggplot2)
ggplot(importance_df, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Single Decision Tree Feature Importance")
```


# Random Forest


```{r}
#install.packages("randomForest")
library(randomForest)
head(state_data)
```

```{r}
rf_model <- randomForest(poverty ~ ., data = train_data)
print(rf_model)
```


```{r}
# Make predictions (this part is now correct!)
predictions <- predict(rf_model, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

```{r}
importance(rf_model)
varImpPlot(rf_model)
```


```{r}
rf_tuned <- randomForest(poverty ~ ., data = train_data, ntree = 500, mtry = 10)
print(rf_tuned)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(rf_tuned, test_data)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")

rss_value <- sum((test_data$poverty - predictions)^2)
cat("RSS:", rss_value, "\n")
```

```{r}
importance(rf_tuned)
varImpPlot(rf_tuned)
```


# Support Vector Regression


```{r}
# Install packages if you haven't already
#install.packages("e1071")

# Load libraries
library(e1071)
```

```{r}
# Train the SVR model
set.seed(1)

svr_model <- svm(poverty ~ ., 
                 data = train_data, 
                 type = "eps-regression", 
                 kernel = "radial",
                 cost = 15, 
                 epsilon = 0.25)

# View model summary details
summary(svr_model)
```

```{r}
# Predict on test data
predicted_Y <- predict(svr_model, test_data)

# Calculate Root Mean Squared Error (RMSE)
error <- test_data$poverty - predicted_Y
rmse <- sqrt(mean(error^2))

# Print performance metric
print(paste("Test RMSE:", round(rmse, 4)))

rss_value <- sum((predicted_Y - test_data$poverty)^2)
cat("RSS:", rss_value, "\n")

# Calculate R-squared (variance explained)
r2_val <- R2(predicted_Y, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# Calculate Mean Absolute Error
mae_val <- MAE(predicted_Y, test_data$poverty)
cat("MAE:", mae_val, "\n")
```


```{r}
ggplot(test_data, aes(x = predicted_Y, y = poverty)) +
  geom_point(color = "red", size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "dashed", size = 1) +
  labs(
    title = "SVR Model Performance (All Predictors)",
    subtitle = "Dashed line represents perfect prediction",
    x = "Predicted Poverty Rate (%)",
    y = "Actual Poverty Rate (%)"
  ) +
  theme_minimal()
```


```{r}
set.seed(3)

# significance of each predictor in the SVR model using permutation importance
# Install and load the vip package
if (!requireNamespace("vip", quietly = TRUE)) install.packages("vip")
library(vip)

# Filter out the prediction column so it doesn't look at its own outputs
importance_data <- test_data[, !names(test_data) %in% c("Predicted_Y")]

# Calculate and plot variable importance
vip(
  object = svr_model, 
  method = "permute", 
  target = "poverty", 
  metric = "rmse",
  pred_wrapper = predict, 
  train = importance_data,
  geom = "col",
  aesthetics = list(fill = "steelblue")
) +
  labs(
    title = "SVR Feature Importance (Permutation)",
    subtitle = "Shows how much each feature reduces RMSE when kept intact",
    x = "Predictors",
    y = "Importance"
  ) +
  theme_minimal()
```


# Comparison Script

```{r}
rf <- rf_tuned
svr <- svr_model
lasso <- fit.min
ols <- fit1
dt <- tree_model
nn <- load_model_native("/Users/winniex/Desktop/WDS_2026_AI_HOME/Project/Models/nn_model_l2_poverty_prediction.keras")
```

## Model Performance Comparison

```{r}
# Create data frame with model results
model_results <- data.frame(
  Model = c("OLS", "LASSO", "Decision Tree", "Random Forest", "SVR", "Deep Neural Net"),
  RSS = c("8.34", "4.92", "20.6", "1.31", "0.506", "13.456"), 
  R_squared = c("0.89", "0.918", "0.418", "0.964", "0.985", "0.803"),
  RMSE = c("1.02", "0.784", "1.61", "0.405", "0.6094", "1.106"),
  MAE = c("0.771", "0.593", "1.21", "0.319", "0.506", "0.956")  
)

# Bar chart for RMSE
ggplot(model_results, aes(x = reorder(Model, RMSE), y = RMSE, fill = Model)) +
  geom_col() +
  coord_flip() +
  labs(title = "Model Performance Comparison",
       subtitle = "Lower RMSE = Better Performance",
       x = "Model", y = "Root Mean Squared Error (RMSE)") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Bar chart for RSS
ggplot(model_results, aes(x = reorder(Model, RSS), y = RSS, fill = Model)) +
  geom_col() +
  coord_flip() +
  labs(title = "Model Performance Comparison",
       subtitle = "Lower RSS = Better Performance",
       x = "Model", y = "Residual Sum of Squares (RSS)") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Bar chart for R_squared
ggplot(model_results, aes(x = reorder(Model, R_squared), y = R_squared, fill = Model)) +
  geom_col() +
  coord_flip() +
  labs(title = "Model Performance Comparison",
       subtitle = "Higher R² = Better Performance",
       x = "Model", y = "R-squared") +
  theme_minimal() +
  theme(legend.position = "none")
```

```{r}
# Bar chart for MAE
ggplot(model_results, aes(x = reorder(Model, MAE), y = MAE, fill = Model)) +
  geom_col() +
  coord_flip() +
  labs(title = "Model Performance Comparison",
       subtitle = "Lower MAE = Better Performance",
       x = "Model", y = "Mean Absolute Error (MAE)") +
  theme_minimal() +
  theme(legend.position = "none")
```

---

## Section: Data Sources (1 graph)

### Graph: Variable Categories Pie Chart

```{r}
# Create category counts
categories <- data.frame(
  Category = c("Demographic", "Economic", "Policy", "Education", "Social", "Infrastructure"),
  Count = c(11, 11, 3, 2, 2, 1)
)

ggplot(categories, aes(x = "", y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Distribution of Predictor Variables by Category",
       subtitle = "Total: 30 predictors") +
  theme_void() +
  scale_fill_brewer(palette = "Set3")
```

---

## Section: EDA - Distribution Analysis (3 graphs)

### Graph: Poverty Boxplot
- **Filename:** `poverty_boxplot.png`
- **Type:** Boxplot
- **Purpose:** Show quartiles and outliers
- **Placement:** Right side of slide
- **R Code:**
```{r}
pov_box <- ggplot(state_data, aes(y = poverty, x = "")) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 3) +
  coord_flip() +
  labs(title = "Poverty Rate Distribution",
       subtitle = "Boxplot showing quartiles and outliers",
       y = "Poverty Rate (%)", x = "") +
  theme_minimal() +
  theme(axis.text.y = element_blank())

pov_box
```

### Graph: Predictor Distributions Grid
- **Filename:** `predictor_distributions_grid.png`
- **Type:** Faceted histograms (small multiples)
- **Purpose:** Show distribution of all 30 predictors
- **Placement:** Full slide (may need 2 slides)
- **R Code:**
```{r}
# Reshape data to long format
data_long <- state_data %>%
  select(-state) %>%  # Remove state names
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create faceted histograms
pdg <- ggplot(data_long, aes(x = value)) +
  geom_histogram(bins = 10, fill = "steelblue", alpha = 0.7) +
  facet_wrap(~ variable, scales = "free", ncol = 5) +
  labs(title = "Distribution of All Predictor Variables",
       x = "Value", y = "Count") +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))

pdg
```

