---
title: "Final Project"
author:
- Winnie
- June
- Helen
- Akshaj
- Lilly
output:
  html_document:
    code_folding: show
    highlight: haddock
    number_sections: yes
    theme: lumen
    toc: yes
    toc_depth: 4
    toc_float: yes
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
    toc_depth: '4'
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=8, fig.height=5, warning=FALSE, message=FALSE)
options(scipen = 0, digits = 3)
```


```{r load-packages}
# Load packages
if(!require('pacman')) install.packages('pacman')
pacman::p_load(tidyverse, readxl, skimr, corrplot, GGally, gridExtra, rpart, rpart.plot, caret, vip, randomForest, e1071)
```


**1. Goal of study**. Restate the goal of study and importance of your project. 

The goal of our project is to analyze the predictors that lead to greater poverty rates. Our project’s model is based on the idea that factors such as Education costs, per capita personal income, population density, unemployment rate and more, determine poverty. Our regression model, along with our analysis, should give the class and our team members a much greater understanding of what factors impact the different communities around the world which suffer from substantial impoverished conditions, ultimately leading us to draw conclusions based upon our findings. 


**2. Data:**

i) What is your data source?

Our data sources are from US department of Labor, US Bureau of Labor Statistics, US Census Bureau, and other official government sources.

ii) Describe how you acquire the data. (You can provide another set of code).

We manually downloaded datasources to assemble a singular dataset for my needs.

iii) Describe how you clean the data (e.g. missing value, errors, outliners).

We didn't have much to do with cleaning the data as we used a standardized government format/official format to collect our data, so there were wasn't much to clean.

iv) Is your data representative enough for your goal of study? Could there be any bias? How will you address the bias if any?

Yes, our data is representative enough due to the wide range of data that we are using in our study. However, aas all of these data sources are from the US government, there might be some bias when it comes to our data.


**3. EDA**

As the first step of our analysis, we need to make sense out of the data by EDA. 
Please provide a complete EDA of your data using relevant plots (with title, labels, colors, legends) and descriptions.
What have you found so far? 


# Research Question


**What state-level variables most strongly predict higher poverty rates across the United States?**


- Unit of analysis: 50 U.S. states
- Time period: 2023
- Outcome: Poverty rate (%)
- Predictors: Economic, social, and policy variables


---


# Load Data


```{r load-data}
library(dplyr)

# Load your cleaned dataset
# Update the file path to match your data location
state_data <- read.csv("/Users/winniex/Desktop/WDS/Project/dataset - Sheet1 (2).csv", sep=",", header=T, as.is=T)


# Show first few rows
head(state_data)
```


```{r data-structure}
# Data structure
glimpse(state_data)
```


---


# Data Quality Check
```{r}
str(state_data)
```

```{r}
summary(state_data)
```

```{r data-summary}
# Summary statistics for all variables
skim(state_data)
```


```{r missing-values}
# Check for missing values
missing_counts <- colSums(is.na(state_data))
missing_counts[missing_counts > 0]
```


```{r basic-stats}
# Basic dataset info
cat("Number of states:", nrow(state_data), "\n")
cat("Number of variables:", ncol(state_data), "\n")
```

```{r}
state_data <- state_data %>%
  rename("state"="State")

```

---


# Outcome Variable: Poverty Rate


```{r poverty-summary}
# Summary statistics
summary(state_data$poverty)
```


```{r poverty-distribution}
# Histogram
ggplot(state_data, aes(x = poverty)) +
 geom_histogram(bins = 10, fill = "steelblue", alpha = 0.7) +
 geom_vline(xintercept = mean(state_data$poverty),
            color = "red", linetype = "dashed", size = 1) +
 labs(title = "Distribution of Poverty Rates Across States",
      subtitle = "Red line = mean",
      x = "Poverty Rate (%)",
      y = "Count") +
 theme_minimal()
```

```{r}
library(usmap)
#names(data.s)  # hist(data.s$crime.rate)
#levels(data.s$state)

stat.poverty.plot <- plot_usmap(regions = "state",                   
                              #regions = "counties", for county level summary
    data = state_data, 
    values = "poverty", color = "black") + 
    labs(title = "State Poverty Rate", subtitle = "Continental US States") +
    theme(legend.position = "right")

stat.poverty.plot
```

---

# Correlation Analysis

```{r}
# Select only numeric variables
numeric_vars <- state_data %>%
 select(where(is.numeric)) 


# Calculate correlation matrix
cor_matrix <- cor(numeric_vars, use = "complete.obs")


# Visualize correlation matrix
corrplot(cor_matrix,
        method = "color",
        type = "upper",
        tl.col = "black",
        title = "Correlation Matrix")
```

```{r}
# Extract correlations with poverty rate
cor_with_poverty <- cor_matrix["poverty", ] %>%
 sort(decreasing = TRUE)


# Display all correlations with poverty
cat("Correlations with poverty rate (sorted):\n")
print(round(cor_with_poverty, 3))
```



```{r correlation-table}
# Top 10 correlations (absolute value)
top_cors <- data.frame(
 Variable = names(cor_with_poverty)[-1],  # exclude poverty itself
 Correlation = cor_with_poverty[-1]
) %>%
 arrange(desc(abs(Correlation))) %>%
 slice(1:10)


cat("\nTop 10 predictors by absolute correlation:\n")
print(top_cors)
```

---

# Scatterplots: Top Predictors


```{r identify-top-predictors}
# Get top 4 predictors (by absolute correlation)
top_4_vars <- top_cors %>%
 slice(1:4) %>%
 pull(Variable)


cat("Top 4 predictors:\n")
print(top_4_vars)
```


```{r scatterplots}
# Create scatterplots for top 4 predictors
plot_list <- lapply(top_4_vars, function(var) {
 ggplot(state_data, aes_string(x = var, y = "poverty")) +
   geom_point(color = "steelblue", size = 2.5, alpha = 0.6) +
   geom_smooth(method = "lm", se = TRUE, color = "red", size = 1) +
   labs(title = paste("Poverty vs", var),
        x = var,
        y = "Poverty Rate (%)") +
   theme_minimal()
})


# Display in grid
grid.arrange(grobs = plot_list, ncol = 2)
```


---


# Pairs Plot


```{r pairs-plot}
# Select poverty rate + top 4 predictors
pairs_data <- state_data %>%
 select(poverty, all_of(top_4_vars))


# Create pairs plot
ggpairs(pairs_data,
       title = "Pairs Plot: Poverty Rate and Top 4 Predictors",
       lower = list(continuous = "smooth"),
       upper = list(continuous = "cor"))
```


---


# Distribution of Key Predictors


```{r predictor-distributions}
state_data <- state_data %>%
  mutate()

# Histograms for top 4 predictors
plot_list_hist <- lapply(top_4_vars, function(var) {
 ggplot(state_data, aes_string(x = var)) +
   geom_histogram(bins = 20, fill = "steelblue", alpha = 0.7) +
   labs(title = var, x = "", y = "Count") +
   theme_minimal()
})


# Display
grid.arrange(grobs = plot_list_hist, ncol = 2)
```


---


# Summary Statistics


```{r summary-stats}
# Summary for poverty rate and top predictors
summary_vars <- state_data %>%
 select(poverty, all_of(top_4_vars))


cat("Summary statistics for key variables:\n")
summary(summary_vars)
```


```{r correlation-strength}
# Categorize correlation strength
top_cors_categorized <- top_cors %>%
 mutate(
   Strength = case_when(
     abs(Correlation) > 0.7 ~ "Strong",
     abs(Correlation) > 0.4 ~ "Moderate",
     abs(Correlation) > 0.2 ~ "Weak",
     TRUE ~ "Very Weak"
   ),
   Direction = ifelse(Correlation > 0, "Positive", "Negative")
 )


cat("\nCorrelation strength and direction:\n")
print(top_cors_categorized)
```

---

# Key Findings


```{r key-findings}
# Calculate key metrics
mean_poverty <- mean(state_data$poverty)
sd_poverty <- sd(state_data$poverty)
min_poverty <- min(state_data$poverty)
max_poverty <- max(state_data$poverty)


cat("=== KEY FINDINGS ===\n\n")
cat("Dataset:\n")
cat("  - States:", nrow(state_data), "\n")
cat("  - Variables:", ncol(state_data), "\n\n")


cat("Poverty Rate:\n")
cat("  - Mean:", round(mean_poverty, 2), "%\n")
cat("  - SD:", round(sd_poverty, 2), "%\n")
cat("  - Range:", round(min_poverty, 2), "% to", round(max_poverty, 2), "%\n\n")


cat("Strongest Predictors:\n")
for(i in 1:4) {
 cat("  ", i, ".", top_cors_categorized$Variable[i],
     "(r =", round(top_cors_categorized$Correlation[i], 3), ")\n")
}


cat("\nNext Steps:\n")
cat("  - Build OLS regression model\n")
cat("  - Run LASSO feature selection\n")
cat("  - Check for multicollinearity (VIF)\n")
cat("  - Consider log transformations for skewed variables\n")
```

**4. Preliminary analysis**

With all the techniques we have learned so far, let's try to build models for your goal of study. 
Please summarize what you have learned from the model. Do the results make sense? 

# Linear Regression

```{r}
fit1 <- lm(poverty ~ .-state, state_data)

summary(fit1)
```

# Relex LASSO

```{r}
library(glmnet)

y <- state_data[,31] # extract Y
x <- model.matrix(poverty~.-state, data=state_data)

set.seed(10)

lasso_cv <- cv.glmnet(x, y, alpha = 1, nfolds=10)

coef(lasso_cv, s = "lambda.1se")
```

```{r}
coef(lasso_cv, s = "lambda.min")
```

```{r}
plot(lasso_cv)
```

```{r}
library(coefplot)

#lambda.min
coef.min <- extract.coef(lasso_cv, lambda="lambda.min")
coef.min
```

```{r}
var_min <- rownames(coef.min)[-1]
min_data <- state_data[, c("poverty", var_min)]
min_data
```

```{r}
#lambda.1se
coef.1se <- extract.coef(lasso_cv, lambda="lambda.1se")
coef.1se
```

```{r}
var_1se <- rownames(coef.1se)[-1]
se_data <- state_data[, c("poverty", var_1se)]
se_data
```

```{r}
#min
fit.min <- lm(poverty ~ ., min_data)
summary(fit.min)
```

```{r}
#1se
fit.1se <- lm(poverty ~ ., se_data)
summary(fit.1se)
```

# Decision Tree

```{r}
#install.packages("rpart")
#install.packages("rpart.plot")
#install.packages("caret")

#Load Data and Split into Train/Test Sets
library(rpart)
library(rpart.plot)
library(caret)

data(iris)
set.seed(123)
train_index <- createDataPartition(state_data$poverty, p = 0.8, list = FALSE)
train_data <- state_data[train_index, ]
test_data <- state_data[-train_index, ]

train_data$state <- NULL
test_data$state  <- NULL
```

```{r}
#Train the Model
tree_model <- rpart(poverty ~ ., 
                    data = train_data, 
                    method = "anova",  # for regression
                    control = rpart.control(minsplit = 10, cp = 0.01))
```

```{r}
#Visualize the Tree
rpart.plot(tree_model, box.palette = "auto", nn = TRUE)
```

```{r}
# Make predictions (this part is now correct!)
predictions <- predict(tree_model, test_data, type = "vector")

# Evaluate Regression performance instead of using a Confusion Matrix
library(caret)

# 1. Calculate Root Mean Squared Error
rmse_val <- RMSE(predictions, test_data$poverty)
cat("RMSE:", rmse_val, "\n")

# 2. Calculate R-squared (variance explained)
r2_val <- R2(predictions, test_data$poverty)
cat("R-squared:", r2_val, "\n")

# 3. Calculate Mean Absolute Error
mae_val <- MAE(predictions, test_data$poverty)
cat("MAE:", mae_val, "\n")
```

```{r}
#Prune the Tree
printcp(tree_model)  # Identify optimal cp (xerror is minimized)
optimal_cp <- tree_model$cptable[which.min(tree_model$cptable[, "xerror"]), "CP"]
pruned_tree <- prune(tree_model, cp = optimal_cp)
rpart.plot(pruned_tree)
```

```{r}
# Tuning Hyperparameters
control <- trainControl(method = "cv", number = 10)  # 10-fold cross-validation
tuned_tree <- train(poverty ~ ., 
                    data = state_data, 
                    method = "rpart",
                    trControl = control,
                    tuneGrid = expand.grid(cp = seq(0.01, 0.1, 0.01)))
print(tuned_tree)
```

```{r}
# Original Model

# 1. Extract the feature importance
# Decision trees store importance directly inside the model object
tree_importance <- tree_model$variable.importance
print(tree_importance)

# 2. Convert to a data frame and plot (same logic as before)
importance_df <- data.frame(
  Feature = names(tree_importance),
  Importance = as.numeric(tree_importance)
)

library(ggplot2)
ggplot(importance_df, aes(x = reorder(Feature, Importance), y = Importance)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  coord_flip() +
  theme_minimal() +
  labs(title = "Single Decision Tree Feature Importance")
```


# Random Forest



```{r}
#install.packages("randomForest")
library(randomForest)
head(state_data)
```

```{r}
set.seed(42)

trainIndex <- sample(1:nrow(state_data), 0.8 * nrow(state_data))

trainData <- state_data[trainIndex, ]
testData <- state_data[-trainIndex, ]

trainData$state <- NULL
testData$state  <- NULL
```
```{r}
rf_model <- randomForest(poverty ~ ., data = trainData)
print(rf_model)
```

```{r}
predictions <- predict(rf_model, testData)
MAE(predictions, testData$poverty)

rf_tuned <- randomForest(poverty ~ ., data = trainData, ntree = 500, mtry = 10)
print(rf_tuned)
```

```{r}
importance(rf_model)
varImpPlot(rf_model)
```

```{r}
# You must set importance = TRUE to calculate the advanced metrics
rf_model <- randomForest(mpg ~ ., data = mtcars, importance = TRUE)

# Extract the importance scores
# This is your equivalent to extract.coef()
importance_scores <- importance(rf_model)
print(importance_scores)

```


# Support Vertor Regression

```{r}
# Install packages if you haven't already
#install.packages("e1071")

# Load libraries
library(e1071)
```

```{r}
# Train the SVR model
set.seed(1)

svr_model <- svm(poverty ~ ., 
                 data = trainData, 
                 type = "eps-regression", 
                 kernel = "radial",
                 cost = 15, 
                 epsilon = 0.25)

# View model summary details
summary(svr_model)
```

```{r}
# Predict on test data
testData$Predicted_Y <- predict(svr_model, testData)

# Calculate Root Mean Squared Error (RMSE)
error <- testData$poverty - testData$Predicted_Y
rmse <- sqrt(mean(error^2))

# Print performance metric
print(paste("Test RMSE:", round(rmse, 4)))
```

```{r}
my_features <- "unemploy_rate"

ggplot(testData) +
  geom_point(aes(x = .data[[my_features]], y = poverty), color = "red", size = 2) +
  geom_line(aes(x = .data[[my_features]], y = Predicted_Y), color = "blue", size = 1) +
  labs(
    title = "Support Vector Regression (Radial Kernel)",
    x = my_features,
    y = "Poverty Rate (%)"
  ) +
  theme_minimal()
```

```{r}
ggplot(testData, aes(x = Predicted_Y, y = poverty)) +
  geom_point(color = "red", size = 2.5) +
  geom_abline(intercept = 0, slope = 1, color = "blue", linetype = "dashed", size = 1) +
  labs(
    title = "SVR Model Performance (All Predictors)",
    subtitle = "Dashed line represents perfect prediction",
    x = "Predicted Poverty Rate (%)",
    y = "Actual Poverty Rate (%)"
  ) +
  theme_minimal()
```

```{r}
set.seed(1)

# significance of each predictor in the SVR model using permutation importance
# Install and load the vip package
if (!requireNamespace("vip", quietly = TRUE)) install.packages("vip")
library(vip)

# Filter out the prediction column so it doesn't look at its own outputs
importance_data <- testData[, !names(testData) %in% c("Predicted_Y")]

# Calculate and plot variable importance
vip(
  object = svr_model, 
  method = "permute", 
  target = "poverty", 
  metric = "rmse",
  pred_wrapper = predict, 
  train = importance_data,
  geom = "col",
  aesthetics = list(fill = "steelblue")
) +
  labs(
    title = "SVR Feature Importance (Permutation)",
    subtitle = "Shows how much each feature reduces RMSE when kept intact",
    x = "Predictors",
    y = "Importance"
  ) +
  theme_minimal()
```


# Comparison Script

```{r}
rf <- rf_model
svr <- svr_model
lasso <- lasso_cv
ols <- fit1
dt <- tree_model
```


## Graph 1: Model Performance Comparison
- **Filename:** `presentation_overview_chart.png`
- **Type:** Bar chart
- **Purpose:** Compare RMSE or R² across all 6 models (OLS, LASSO, Decision Tree, Random Forest, SVR, DNN)
- **Placement:** Right side of executive summary slide
- **R Code:**
```{r}
library(ggplot2)
library(dplyr)

# Create data frame with model results
model_results <- data.frame(
  Model = c("OLS", "LASSO", "Decision Tree", "Random Forest", "SVR", "Deep Neural Net"),
  RMSE = c("0.623", "1.456", "1.67", "1.311", "1.367", ""),  # Insert your actual values
  R_squared = c("0.931", "0.942", "0.418", "", "", "")  # Insert your actual values
)

# Bar chart for RMSE
ggplot(model_results, aes(x = reorder(Model, RMSE), y = RMSE, fill = Model)) +
  geom_col() +
  coord_flip() +
  labs(title = "Model Performance Comparison",
       subtitle = "Lower RMSE = Better Performance",
       x = "Model", y = "Root Mean Squared Error (RMSE)") +
  theme_minimal() +
  theme(legend.position = "none")

ggsave("output/figures/presentation_overview_chart.png", width = 8, height = 5, dpi = 300)
```

---

## Section 2: Research Question (1 graph)

### Graph 2: U.S. Poverty Map
- **Filename:** `us_poverty_map.png`
- **Type:** Choropleth map
- **Purpose:** Show geographic distribution of poverty rates by state
- **Placement:** Bottom of slide
- **R Code:**
```r
library(maps)
library(viridis)
library(ggplot2)

# Load state data
states_map <- map_data("state")

# Merge with your poverty data
# (Assuming you have a data frame called 'state_data' with columns: state, poverty)
state_data$region <- tolower(state_data$state)

map_data <- states_map %>%
  left_join(state_data, by = "region")

# Create choropleth
ggplot(map_data, aes(x = long, y = lat, group = group, fill = poverty)) +
  geom_polygon(color = "white", size = 0.2) +
  scale_fill_viridis(option = "plasma", name = "Poverty Rate (%)") +
  coord_map("albers", lat0 = 39, lat1 = 45) +
  labs(title = "State Poverty Rates Across the United States (2023)") +
  theme_void() +
  theme(legend.position = "right")

ggsave("output/figures/us_poverty_map.png", width = 10, height = 6, dpi = 300)
```

---

## Section 3: Data Sources (1 graph)

### Graph 3: Variable Categories Pie Chart
- **Filename:** `variable_categories_pie.png`
- **Type:** Pie/donut chart
- **Purpose:** Show distribution of 30 variables by category
- **Placement:** Right side of slide
- **R Code:**
```r
# Create category counts
categories <- data.frame(
  Category = c("Demographic", "Economic", "Policy", "Education", "Social", "Infrastructure"),
  Count = c(11, 11, 3, 2, 2, 1)
)

ggplot(categories, aes(x = "", y = Count, fill = Category)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Distribution of Predictor Variables by Category",
       subtitle = "Total: 30 predictors") +
  theme_void() +
  scale_fill_brewer(palette = "Set3")

ggsave("output/figures/variable_categories_pie.png", width = 6, height = 6, dpi = 300)
```

---

## Section 4: Dataset Overview (1 graph/table)

### Graph 4: Summary Statistics Table
- **Filename:** `summary_statistics_table.png`
- **Type:** Formatted table
- **Purpose:** Show descriptive statistics for key variables
- **Placement:** Center of slide
- **R Code:**
```r
library(knitr)
library(kableExtra)

# Create summary statistics
summary_stats <- data %>%
  select(poverty, log_income, unemploy_rate, crime) %>%
  summarise(across(everything(),
                   list(Mean = mean,
                        Median = median,
                        SD = sd,
                        Min = min,
                        Max = max))) %>%
  pivot_longer(everything(),
               names_to = c("Variable", "Statistic"),
               names_sep = "_") %>%
  pivot_wider(names_from = Statistic, values_from = value)

# Create formatted table
kable(summary_stats, format = "html", digits = 2,
      caption = "Summary Statistics for Key Variables") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) %>%
  save_kable("output/figures/summary_statistics_table.png")
```

---

## Section 5: EDA - Distribution Analysis (3 graphs)

### Graph 5: Poverty Distribution Histogram
- **Filename:** `poverty_distribution_histogram.png`
- **Type:** Histogram with normal curve overlay
- **Purpose:** Show distribution of outcome variable
- **Placement:** Left side of slide
- **R Code:**
```r
ggplot(data, aes(x = poverty)) +
  geom_histogram(aes(y = ..density..), bins = 15, fill = "steelblue", alpha = 0.7) +
  stat_function(fun = dnorm,
                args = list(mean = mean(data$poverty), sd = sd(data$poverty)),
                color = "red", size = 1) +
  labs(title = "Distribution of State Poverty Rates",
       x = "Poverty Rate (%)", y = "Density") +
  theme_minimal()

ggsave("output/figures/poverty_distribution_histogram.png", width = 6, height = 4, dpi = 300)
```

### Graph 6: Poverty Boxplot
- **Filename:** `poverty_boxplot.png`
- **Type:** Boxplot
- **Purpose:** Show quartiles and outliers
- **Placement:** Right side of slide
- **R Code:**
```r
ggplot(data, aes(y = poverty, x = "")) +
  geom_boxplot(fill = "lightblue", outlier.color = "red", outlier.size = 3) +
  coord_flip() +
  labs(title = "Poverty Rate Distribution",
       subtitle = "Boxplot showing quartiles and outliers",
       y = "Poverty Rate (%)", x = "") +
  theme_minimal() +
  theme(axis.text.y = element_blank())

ggsave("output/figures/poverty_boxplot.png", width = 6, height = 3, dpi = 300)
```

### Graph 7: Predictor Distributions Grid
- **Filename:** `predictor_distributions_grid.png`
- **Type:** Faceted histograms (small multiples)
- **Purpose:** Show distribution of all 30 predictors
- **Placement:** Full slide (may need 2 slides)
- **R Code:**
```r
# Reshape data to long format
data_long <- data %>%
  select(-state) %>%  # Remove state names
  pivot_longer(everything(), names_to = "variable", values_to = "value")

# Create faceted histograms
ggplot(data_long, aes(x = value)) +
  geom_histogram(bins = 10, fill = "steelblue", alpha = 0.7) +
  facet_wrap(~ variable, scales = "free", ncol = 5) +
  labs(title = "Distribution of All Predictor Variables",
       x = "Value", y = "Count") +
  theme_minimal() +
  theme(strip.text = element_text(size = 8))

ggsave("output/figures/predictor_distributions_grid.png", width = 15, height = 10, dpi = 300)
```





 

---

# Cross-Validation & Model Evaluation

This section performs comprehensive cross-validation for all models to ensure robust performance evaluation and prevent overfitting.

## Setup: Train/Test Split & CV Configuration

```{r cv-setup}
# Set seed for reproducibility
set.seed(42)

# Create output directories
dir.create("../output/figures", recursive = TRUE, showWarnings = FALSE)
dir.create("../output/tables", recursive = TRUE, showWarnings = FALSE)

# Prepare data (remove state names)
data_for_modeling <- state_data %>% select(-state)

# Create train/test split (80/20)
train_index <- createDataPartition(data_for_modeling$poverty, p = 0.8, list = FALSE)

train_data_cv <- data_for_modeling[train_index, ]
test_data_cv <- data_for_modeling[-train_index, ]

# Store test state names for later
test_states_cv <- state_data$state[-train_index]

cat("Training set:", nrow(train_data_cv), "observations\n")
cat("Test set:", nrow(test_data_cv), "observations\n")

# Prepare X and y matrices for LASSO/GLMnet
X_train_cv <- as.matrix(train_data_cv %>% select(-poverty))
y_train_cv <- train_data_cv$poverty

X_test_cv <- as.matrix(test_data_cv %>% select(-poverty))
y_test_cv <- test_data_cv$poverty

# Define 5-fold cross-validation control
cv_control <- trainControl(
  method = "cv",           # Cross-validation
  number = 5,              # 5 folds
  savePredictions = "final",
  verboseIter = FALSE
)
```

---

## Model 1: OLS Regression with Cross-Validation

```{r ols-cv}
cat("\n========== OLS REGRESSION ==========\n")

# Train OLS with 5-fold CV
ols_cv <- train(
  poverty ~ .,
  data = train_data_cv,
  method = "lm",
  trControl = cv_control,
  metric = "RMSE"
)

# Extract CV results
ols_cv_results <- ols_cv$resample
ols_cv_rmse <- mean(ols_cv_results$RMSE)
ols_cv_rsq <- mean(ols_cv_results$Rsquared)
ols_cv_rmse_sd <- sd(ols_cv_results$RMSE)

# Fit final model on full training set
ols_final <- lm(poverty ~ ., data = train_data_cv)

# Training set predictions
ols_train_pred <- predict(ols_final, train_data_cv)
ols_train_rmse <- sqrt(mean((train_data_cv$poverty - ols_train_pred)^2))
ols_train_rsq <- cor(train_data_cv$poverty, ols_train_pred)^2

# Test set predictions
ols_test_pred <- predict(ols_final, test_data_cv)
ols_test_rmse <- sqrt(mean((y_test_cv - ols_test_pred)^2))
ols_test_rsq <- cor(y_test_cv, ols_test_pred)^2
ols_test_mae <- mean(abs(y_test_cv - ols_test_pred))

# Display results
cat("Cross-Validation RMSE:", round(ols_cv_rmse, 3), "±", round(ols_cv_rmse_sd, 3), "\n")
cat("Train RMSE:", round(ols_train_rmse, 3), "| R²:", round(ols_train_rsq, 3), "\n")
cat("Test RMSE:", round(ols_test_rmse, 3), "| R²:", round(ols_test_rsq, 3), "\n")
cat("Test MAE:", round(ols_test_mae, 3), "\n")
```

### OLS Model Summary

```{r ols-summary}
summary(ols_final)
```

---

## Model 2: LASSO with Cross-Validation

```{r lasso-cv}
cat("\n========== LASSO REGRESSION ==========\n")

# Set up lambda sequence
lambda_seq <- 10^seq(-2, 2, length = 100)

# Train LASSO with CV
lasso_cv_model <- train(
  poverty ~ .,
  data = train_data_cv,
  method = "glmnet",
  trControl = cv_control,
  tuneGrid = expand.grid(alpha = 1, lambda = lambda_seq),
  metric = "RMSE"
)

# Extract CV results for optimal lambda
lasso_cv_results <- lasso_cv_model$resample %>%
  filter(lambda == lasso_cv_model$bestTune$lambda)
lasso_cv_rmse <- mean(lasso_cv_results$RMSE)
lasso_cv_rsq <- mean(lasso_cv_results$Rsquared)
lasso_cv_rmse_sd <- sd(lasso_cv_results$RMSE)

# Optimal lambda
optimal_lambda <- lasso_cv_model$bestTune$lambda

# Fit final LASSO model
lasso_final <- glmnet(X_train_cv, y_train_cv, alpha = 1, lambda = optimal_lambda)

# Training predictions
lasso_train_pred <- predict(lasso_final, X_train_cv)[,1]
lasso_train_rmse <- sqrt(mean((y_train_cv - lasso_train_pred)^2))
lasso_train_rsq <- cor(y_train_cv, lasso_train_pred)^2

# Test predictions
lasso_test_pred <- predict(lasso_final, X_test_cv)[,1]
lasso_test_rmse <- sqrt(mean((y_test_cv - lasso_test_pred)^2))
lasso_test_rsq <- cor(y_test_cv, lasso_test_pred)^2
lasso_test_mae <- mean(abs(y_test_cv - lasso_test_pred))

# Count non-zero coefficients
lasso_coefs <- coef(lasso_final)
n_features_selected <- sum(lasso_coefs[-1] != 0)

cat("Optimal lambda:", round(optimal_lambda, 4), "\n")
cat("Features selected:", n_features_selected, "out of", ncol(X_train_cv), "\n")
cat("CV RMSE:", round(lasso_cv_rmse, 3), "±", round(lasso_cv_rmse_sd, 3), "\n")
cat("Train RMSE:", round(lasso_train_rmse, 3), "| R²:", round(lasso_train_rsq, 3), "\n")
cat("Test RMSE:", round(lasso_test_rmse, 3), "| R²:", round(lasso_test_rsq, 3), "\n")
cat("Test MAE:", round(lasso_test_mae, 3), "\n")
```

### LASSO Coefficient Path

```{r lasso-path-plot}
# Plot CV curve
plot(lasso_cv_model$finalModel, xvar = "lambda", label = TRUE,
     main = "LASSO Coefficient Paths")
abline(v = log(optimal_lambda), lty = 2, col = "red")
```

---

## Model 3: Decision Tree with Cross-Validation

```{r tree-cv}
cat("\n========== DECISION TREE ==========\n")

# Train Decision Tree with CV
tree_cv_model <- train(
  poverty ~ .,
  data = train_data_cv,
  method = "rpart",
  trControl = cv_control,
  tuneLength = 10,
  metric = "RMSE"
)

# Extract CV results
tree_cv_results <- tree_cv_model$resample
tree_cv_rmse <- mean(tree_cv_results$RMSE)
tree_cv_rsq <- mean(tree_cv_results$Rsquared)
tree_cv_rmse_sd <- sd(tree_cv_results$RMSE)

# Fit final tree
tree_final <- rpart(poverty ~ ., data = train_data_cv,
                    cp = tree_cv_model$bestTune$cp)

# Training predictions
tree_train_pred <- predict(tree_final, train_data_cv)
tree_train_rmse <- sqrt(mean((y_train_cv - tree_train_pred)^2))
tree_train_rsq <- cor(y_train_cv, tree_train_pred)^2

# Test predictions
tree_test_pred <- predict(tree_final, test_data_cv)
tree_test_rmse <- sqrt(mean((y_test_cv - tree_test_pred)^2))
tree_test_rsq <- cor(y_test_cv, tree_test_pred)^2
tree_test_mae <- mean(abs(y_test_cv - tree_test_pred))

cat("Optimal CP:", round(tree_cv_model$bestTune$cp, 4), "\n")
cat("CV RMSE:", round(tree_cv_rmse, 3), "±", round(tree_cv_rmse_sd, 3), "\n")
cat("Train RMSE:", round(tree_train_rmse, 3), "| R²:", round(tree_train_rsq, 3), "\n")
cat("Test RMSE:", round(tree_test_rmse, 3), "| R²:", round(tree_test_rsq, 3), "\n")
cat("Test MAE:", round(tree_test_mae, 3), "\n")
```

---

## Model 4: Random Forest with Cross-Validation

```{r rf-cv, cache=TRUE}
cat("\n========== RANDOM FOREST ==========\n")

# Train Random Forest with CV
# Note: This may take 1-2 minutes
rf_cv_model <- train(
  poverty ~ .,
  data = train_data_cv,
  method = "rf",
  trControl = cv_control,
  tuneGrid = expand.grid(mtry = c(3, 5, 7, 10)),
  ntree = 500,
  importance = TRUE,
  metric = "RMSE"
)

# Extract CV results
rf_cv_results <- rf_cv_model$resample
rf_cv_rmse <- mean(rf_cv_results$RMSE)
rf_cv_rsq <- mean(rf_cv_results$Rsquared)
rf_cv_rmse_sd <- sd(rf_cv_results$RMSE)

# Fit final Random Forest
rf_final <- randomForest(
  poverty ~ .,
  data = train_data_cv,
  ntree = 500,
  mtry = rf_cv_model$bestTune$mtry,
  importance = TRUE
)

# Training predictions
rf_train_pred <- predict(rf_final, train_data_cv)
rf_train_rmse <- sqrt(mean((y_train_cv - rf_train_pred)^2))
rf_train_rsq <- cor(y_train_cv, rf_train_pred)^2

# Test predictions
rf_test_pred <- predict(rf_final, test_data_cv)
rf_test_rmse <- sqrt(mean((y_test_cv - rf_test_pred)^2))
rf_test_rsq <- cor(y_test_cv, rf_test_pred)^2
rf_test_mae <- mean(abs(y_test_cv - rf_test_pred))

cat("Optimal mtry:", rf_cv_model$bestTune$mtry, "\n")
cat("CV RMSE:", round(rf_cv_rmse, 3), "±", round(rf_cv_rmse_sd, 3), "\n")
cat("Train RMSE:", round(rf_train_rmse, 3), "| R²:", round(rf_train_rsq, 3), "\n")
cat("Test RMSE:", round(rf_test_rmse, 3), "| R²:", round(rf_test_rsq, 3), "\n")
cat("Test MAE:", round(rf_test_mae, 3), "\n")
```

### Random Forest Variable Importance

```{r rf-importance-plot}
varImpPlot(rf_final, main = "Random Forest: Variable Importance")
```

---

## Model 5: Support Vector Regression with Cross-Validation

```{r svr-cv}
cat("\n========== SUPPORT VECTOR REGRESSION ==========\n")

# Train SVR with CV (using radial kernel)
svr_cv_model <- train(
  poverty ~ .,
  data = train_data_cv,
  method = "svmRadial",
  trControl = cv_control,
  tuneLength = 5,
  metric = "RMSE",
  preProcess = c("center", "scale")
)

# Extract CV results
svr_cv_results <- svr_cv_model$resample
svr_cv_rmse <- mean(svr_cv_results$RMSE)
svr_cv_rsq <- mean(svr_cv_results$Rsquared)
svr_cv_rmse_sd <- sd(svr_cv_results$RMSE)

# Fit final SVR model
# Scale data for SVR
train_scaled <- scale(train_data_cv %>% select(-poverty))
test_scaled <- scale(test_data_cv %>% select(-poverty),
                     center = attr(train_scaled, "scaled:center"),
                     scale = attr(train_scaled, "scaled:scale"))

svr_final <- svm(
  poverty ~ .,
  data = data.frame(poverty = y_train_cv, train_scaled),
  kernel = "radial",
  cost = svr_cv_model$bestTune$C,
  gamma = svr_cv_model$bestTune$sigma
)

# Training predictions
svr_train_pred <- predict(svr_final, data.frame(train_scaled))
svr_train_rmse <- sqrt(mean((y_train_cv - svr_train_pred)^2))
svr_train_rsq <- cor(y_train_cv, svr_train_pred)^2

# Test predictions
svr_test_pred <- predict(svr_final, data.frame(test_scaled))
svr_test_rmse <- sqrt(mean((y_test_cv - svr_test_pred)^2))
svr_test_rsq <- cor(y_test_cv, svr_test_pred)^2
svr_test_mae <- mean(abs(y_test_cv - svr_test_pred))

cat("Optimal C:", round(svr_cv_model$bestTune$C, 4), "\n")
cat("Optimal sigma:", round(svr_cv_model$bestTune$sigma, 4), "\n")
cat("CV RMSE:", round(svr_cv_rmse, 3), "±", round(svr_cv_rmse_sd, 3), "\n")
cat("Train RMSE:", round(svr_train_rmse, 3), "| R²:", round(svr_train_rsq, 3), "\n")
cat("Test RMSE:", round(svr_test_rmse, 3), "| R²:", round(svr_test_rsq, 3), "\n")
cat("Test MAE:", round(svr_test_mae, 3), "\n")
```

---

# Performance Summary & Comparison

## Overall Performance Table

```{r performance-table}
# Compile all results into one table
performance_summary <- data.frame(
  Model = c("OLS Regression", "LASSO", "Decision Tree", "Random Forest", "SVR"),

  # Cross-Validation Results
  CV_RMSE = c(ols_cv_rmse, lasso_cv_rmse, tree_cv_rmse, rf_cv_rmse, svr_cv_rmse),
  CV_RMSE_SD = c(ols_cv_rmse_sd, lasso_cv_rmse_sd, tree_cv_rmse_sd,
                 rf_cv_rmse_sd, svr_cv_rmse_sd),
  CV_Rsquared = c(ols_cv_rsq, lasso_cv_rsq, tree_cv_rsq, rf_cv_rsq, svr_cv_rsq),

  # Training Set Results
  Train_RMSE = c(ols_train_rmse, lasso_train_rmse, tree_train_rmse,
                 rf_train_rmse, svr_train_rmse),
  Train_Rsquared = c(ols_train_rsq, lasso_train_rsq, tree_train_rsq,
                     rf_train_rsq, svr_train_rsq),

  # Test Set Results
  Test_RMSE = c(ols_test_rmse, lasso_test_rmse, tree_test_rmse,
                rf_test_rmse, svr_test_rmse),
  Test_Rsquared = c(ols_test_rsq, lasso_test_rsq, tree_test_rsq,
                    rf_test_rsq, svr_test_rsq),
  Test_MAE = c(ols_test_mae, lasso_test_mae, tree_test_mae, rf_test_mae, svr_test_mae)
) %>%
  mutate(
    Overfitting = ifelse(Train_RMSE < Test_RMSE - 0.5, "Yes", "No")
  )

# Display formatted table
kable(performance_summary %>%
        mutate(across(where(is.numeric), ~round(., 3))),
      caption = "Model Performance Summary: Cross-Validation and Test Set Metrics",
      col.names = c("Model", "CV RMSE", "CV SD", "CV R²",
                    "Train RMSE", "Train R²", "Test RMSE", "Test R²",
                    "Test MAE", "Overfitting?")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  row_spec(which.min(performance_summary$Test_RMSE),
           bold = TRUE, color = "white", background = "#28a745")
```

```{r save-performance-table}
# Save to CSV
write_csv(performance_summary, "../output/tables/model_performance_summary.csv")
cat("✓ Table saved to output/tables/model_performance_summary.csv\n")
```

---

## Best Model Summary

```{r best-model}
best_idx <- which.min(performance_summary$Test_RMSE)

cat("\n========================================\n")
cat("BEST MODEL (by Test RMSE)\n")
cat("========================================\n")
cat("Model:", performance_summary$Model[best_idx], "\n")
cat("Test RMSE:", round(performance_summary$Test_RMSE[best_idx], 3), "\n")
cat("Test R²:", round(performance_summary$Test_Rsquared[best_idx], 3), "\n")
cat("Test MAE:", round(performance_summary$Test_MAE[best_idx], 3), "\n")
cat("Overfitting:", performance_summary$Overfitting[best_idx], "\n")
cat("========================================\n")
```

---

# Visualization: Model Comparison

## Graph 1: Train vs. Test RMSE Comparison

```{r rmse-comparison-plot, fig.width=10, fig.height=6}
rmse_comparison <- performance_summary %>%
  select(Model, Train_RMSE, Test_RMSE) %>%
  pivot_longer(cols = c(Train_RMSE, Test_RMSE),
               names_to = "Dataset",
               values_to = "RMSE") %>%
  mutate(Dataset = recode(Dataset, "Train_RMSE" = "Train", "Test_RMSE" = "Test"))

ggplot(rmse_comparison, aes(x = Model, y = RMSE, fill = Dataset)) +
  geom_col(position = "dodge", width = 0.7) +
  labs(title = "Model Performance: Train vs. Test RMSE",
       subtitle = "Lower RMSE = Better Performance",
       x = "Model", y = "Root Mean Squared Error (RMSE)",
       fill = "Dataset") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(size = 14, face = "bold")) +
  scale_fill_manual(values = c("Train" = "#4ECDC4", "Test" = "#FF6B6B"))

ggsave("../output/figures/model_comparison_rmse.png", width = 10, height = 6, dpi = 300)
```

## Graph 2: Test R² Comparison (Lollipop Chart)

```{r r2-comparison-plot, fig.width=8, fig.height=6}
ggplot(performance_summary, aes(x = reorder(Model, Test_Rsquared), y = Test_Rsquared)) +
  geom_segment(aes(x = Model, xend = Model, y = 0, yend = Test_Rsquared),
               size = 1.5, color = "gray50") +
  geom_point(size = 5, color = "#1E88E5") +
  coord_flip() +
  labs(title = "Model Performance: R² Values (Test Set)",
       subtitle = "Higher R² = Better Fit",
       x = "Model", y = "R² (Proportion of Variance Explained)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold")) +
  ylim(0, 1) +
  geom_hline(yintercept = 0.7, linetype = "dashed", color = "red", alpha = 0.5)

ggsave("../output/figures/model_comparison_r2.png", width = 8, height = 6, dpi = 300)
```

## Graph 3: Cross-Validation Performance (Boxplot)

```{r cv-boxplot, fig.width=10, fig.height=6}
# Combine CV results from all models
cv_by_fold <- data.frame(
  Model = rep(c("OLS", "LASSO", "Tree", "RF", "SVR"), each = 5),
  Fold = rep(1:5, 5),
  RMSE = c(
    ols_cv_results$RMSE,
    lasso_cv_results$RMSE,
    tree_cv_results$RMSE,
    rf_cv_results$RMSE,
    svr_cv_results$RMSE
  ),
  Rsquared = c(
    ols_cv_results$Rsquared,
    lasso_cv_results$Rsquared,
    tree_cv_results$Rsquared,
    rf_cv_results$Rsquared,
    svr_cv_results$Rsquared
  )
)

ggplot(cv_by_fold, aes(x = reorder(Model, RMSE, FUN = median), y = RMSE, fill = Model)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(width = 0.15, alpha = 0.4, size = 2) +
  labs(title = "Cross-Validation Performance (5-Fold CV)",
       subtitle = "Distribution of RMSE across folds",
       x = "Model", y = "RMSE") +
  theme_minimal() +
  theme(legend.position = "none",
        plot.title = element_text(size = 14, face = "bold")) +
  scale_fill_brewer(palette = "Set2")

ggsave("../output/figures/cv_performance_boxplot.png", width = 10, height = 6, dpi = 300)
```

## Graph 4: Predicted vs. Actual (All Models)

```{r predicted-actual-plot, fig.width=12, fig.height=8}
pred_vs_actual <- data.frame(
  Actual = rep(y_test_cv, 5),
  Predicted = c(ols_test_pred, lasso_test_pred, tree_test_pred,
                rf_test_pred, svr_test_pred),
  Model = rep(c("OLS", "LASSO", "Decision Tree", "Random Forest", "SVR"),
              each = length(y_test_cv))
)

ggplot(pred_vs_actual, aes(x = Actual, y = Predicted)) +
  geom_point(alpha = 0.6, color = "#1E88E5", size = 3) +
  geom_abline(slope = 1, intercept = 0, color = "red", linetype = "dashed", size = 1) +
  facet_wrap(~ Model, ncol = 3) +
  labs(title = "Predicted vs. Actual Poverty Rates (Test Set)",
       subtitle = "Red dashed line = perfect prediction",
       x = "Actual Poverty Rate (%)",
       y = "Predicted Poverty Rate (%)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        strip.text = element_text(size = 10, face = "bold"))

ggsave("../output/figures/predicted_vs_actual_all_models.png",
       width = 12, height = 8, dpi = 300)
```

---

# Feature Importance Comparison

## Random Forest Importance

```{r rf-importance-table}
rf_importance <- importance(rf_final) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  arrange(desc(`%IncMSE`)) %>%
  slice(1:15)

kable(rf_importance,
      caption = "Top 15 Most Important Variables (Random Forest)",
      digits = 3) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Save
write_csv(rf_importance, "../output/tables/rf_variable_importance.csv")
```

## LASSO Selected Features

```{r lasso-features-table}
lasso_coefs_df <- as.matrix(coef(lasso_final)) %>%
  as.data.frame() %>%
  rownames_to_column("Variable") %>%
  rename(Coefficient = s0) %>%
  filter(Variable != "(Intercept)", abs(Coefficient) > 0) %>%
  arrange(desc(abs(Coefficient)))

kable(lasso_coefs_df,
      caption = "LASSO Selected Features (Non-Zero Coefficients)",
      digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Save
write_csv(lasso_coefs_df, "../output/tables/lasso_coefficients.csv")
```

## OLS Significant Predictors

```{r ols-significant-table}
ols_coefs_df <- tidy(ols_final) %>%
  filter(term != "(Intercept)", p.value < 0.05) %>%
  arrange(desc(abs(estimate))) %>%
  select(term, estimate, std.error, statistic, p.value)

kable(ols_coefs_df,
      caption = "OLS Significant Predictors (p < 0.05)",
      digits = 4) %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

# Save
write_csv(ols_coefs_df, "../output/tables/ols_coefficients.csv")
```

---

# Test Set Predictions by State

```{r test-predictions-table}
test_predictions <- data.frame(
  State = test_states_cv,
  Actual = y_test_cv,
  OLS = ols_test_pred,
  LASSO = lasso_test_pred,
  Tree = tree_test_pred,
  RF = rf_test_pred,
  SVR = svr_test_pred
) %>%
  mutate(
    Best_Model = apply(select(., OLS:SVR), 1, function(x) {
      names(which.min(abs(x - Actual)))
    }),
    Best_Prediction = apply(select(., OLS:SVR), 1, function(x) {
      x[which.min(abs(x - Actual))]
    }),
    Best_Error = abs(Actual - Best_Prediction)
  )

kable(test_predictions %>% mutate(across(where(is.numeric), ~round(., 2))),
      caption = "Test Set Predictions by State") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE, font_size = 11)

# Save
write_csv(test_predictions, "../output/tables/test_set_predictions.csv")
```

---

# Save Model Objects

```{r save-models}
# Create models directory
dir.create("../models", showWarnings = FALSE)

# Save all final models
saveRDS(ols_final, "../models/ols_model.rds")
saveRDS(lasso_final, "../models/lasso_model.rds")
saveRDS(tree_final, "../models/tree_model.rds")
saveRDS(rf_final, "../models/rf_model.rds")
saveRDS(svr_final, "../models/svr_model.rds")

# Save train/test split for reproducibility
saveRDS(list(
  train_index = train_index,
  test_states = test_states_cv,
  X_train = X_train_cv,
  y_train = y_train_cv,
  X_test = X_test_cv,
  y_test = y_test_cv
), "../models/train_test_split.rds")

cat("✓ All models saved to models/ directory\n")
```

---

# Summary & Conclusions

```{r final-summary}
cat("\n")
cat(paste(rep("=", 70), collapse = ""), "\n")
cat("CROSS-VALIDATION & MODEL EVALUATION COMPLETE\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")

cat("Files Generated:\n\n")

cat("Tables (5):\n")
cat("  ✓ output/tables/model_performance_summary.csv\n")
cat("  ✓ output/tables/rf_variable_importance.csv\n")
cat("  ✓ output/tables/lasso_coefficients.csv\n")
cat("  ✓ output/tables/ols_coefficients.csv\n")
cat("  ✓ output/tables/test_set_predictions.csv\n\n")

cat("Graphs (4):\n")
cat("  ✓ output/figures/model_comparison_rmse.png\n")
cat("  ✓ output/figures/model_comparison_r2.png\n")
cat("  ✓ output/figures/cv_performance_boxplot.png\n")
cat("  ✓ output/figures/predicted_vs_actual_all_models.png\n\n")

cat("Model Objects (6):\n")
cat("  ✓ models/ols_model.rds\n")
cat("  ✓ models/lasso_model.rds\n")
cat("  ✓ models/tree_model.rds\n")
cat("  ✓ models/rf_model.rds\n")
cat("  ✓ models/svr_model.rds\n")
cat("  ✓ models/train_test_split.rds\n\n")

cat(paste(rep("=", 70), collapse = ""), "\n")
cat("Best Model:", performance_summary$Model[best_idx], "\n")
cat("Test RMSE:", round(performance_summary$Test_RMSE[best_idx], 3), "\n")
cat("Test R²:", round(performance_summary$Test_Rsquared[best_idx], 3), "\n")
cat(paste(rep("=", 70), collapse = ""), "\n\n")
```

**Next Steps:**
1. Review the performance comparison table to identify the best model
2. Examine cross-validation boxplot for model stability
3. Check predicted vs. actual plots for systematic errors
4. Review feature importance to understand key drivers
5. Use these results to update your final presentation report
